Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS

### 2022-04-19 R/CONUS_AOD.Rmd
update report for paper

---
title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
output: html_document
---

@@

## Stratified results for Terra CV dataset
```{r stratified, echo=FALSE}
performance_aod <- function(dt, ...){
  performance_metrics(dt,
                      ground_truth = "AOD_470nm",
                      eo_raw = "MCD19_AOD_470nm",
                      eo_pred = "aodhat")
}
tar_load(initial_cv_l2_terra_conus)
dt <- initial_cv_l2_terra_conus$mDT_wPred
if(!"aodhat" %in% colnames(dt)){
  dt[, aodhat := MCD19_AOD_470nm - diff_AOD_pred]
}

dtwrapper <- function(dt){
  DT::datatable(dt,
                rownames = FALSE,
                fillContainer = FALSE,
                autoHideNavigation = TRUE,
                options = list(pageLength = 50,
                               autoWidth = TRUE))
}
```

### Performance by year
```{r strata_year, echo=FALSE, eval=TRUE}
dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
```

@@

*We can customize however we want*

```{r agreement_plot, warning=FALSE, echo=FALSE}
p1 <- ggplot(dt) +
  aes(AOD_470nm, MCD19_AOD_470nm) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

p2  <- ggplot(dt) +
  aes(AOD_470nm, aodhat) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("Corrected AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

@@

## SHAP interpretation ####
In decreasing order (from Terra collocated cross-validation predictions)
```{r shap_rank, echo=FALSE}
# what do the SHAPs tell us?
tar_load(initial_cv_l2_terra_conus)
shaps <- initial_cv_l2_terra_conus$"shap_score"
# sort by importance
kable(sort(sapply(abs(shaps), mean), decreasing = TRUE), 
      format = "html",
      table.attr = "style='width:40%;'",
      digits = 4, col.names = "Mean Abs SHAP")

>>>>

---
title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
output: html_document
bibliography: references.json
---

@@

## Stratified results for Terra CV dataset

```{r stratified, echo=FALSE}
tar_load(initial_cv_l2_terra_conus)
dt <- initial_cv_l2_terra_conus$mDT_wPred
if(!"aodhat" %in% colnames(dt)){
  dt[, aodhat := MCD19_AOD_470nm - diff_AOD_pred]
}
```

### Performance by year

```{r strata_year, echo=FALSE, eval=TRUE}
dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
```

@@

```{r agreement_plot, warning=FALSE, echo=FALSE}
p1 <- ggplot(dt[AOD_470nm < 1 & MCD19_AOD_470nm < 1]) +
  aes(AOD_470nm, MCD19_AOD_470nm) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1) + 
  labs(caption = paste0("Not shown: ", dt[AOD_470nm > 1 | MCD19_AOD_470nm > 1, .N], " AOD values outside [0, 1]"))

p2  <- ggplot(dt[AOD_470nm < 1 & aodhat > 0 & aodhat < 1]) +
  aes(AOD_470nm, aodhat) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("Corrected AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1) + 
  labs(caption = paste0("Not shown: ", dt[AOD_470nm > 1 | aodhat > 1 | aodhat < 0, .N], " AOD values outside [0, 1]"))

@@

## SHAP interpretation

In decreasing order (from Terra collocated cross-validation predictions)

```{r shap_rank, echo=FALSE}
# what do the SHAPs tell us?
tar_load(initial_cv_l2_terra_conus)
shaps <- initial_cv_l2_terra_conus$"shap_score"
# sort by importance
kable(sort(sapply(abs(shaps), mean), decreasing = TRUE), 
      format = "html",
      table.attr = "style='width:30%;'",
      digits = 3, col.names = "Mean Abs SHAP")
```

*End of report*
